---
title: "Graphics and Text"
author: "JML"
date: "`Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide 
    number_sections: TRUE
    theme: sandstone
    highlight: haddock
    fig_caption: FALSE
---
```{r, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(sysfonts)
library(knitr)
library(ggforce)
```

# Font-related Packages  

  *  `sysfont` package. Can pull in [Google fonts](https://fonts.google.com/)    
  *  `extrafont` package. Very useful for TTF in PDF  
  *  `showtext` package. Flexible for non-standard fonts in R graphs. Doesn't work with RStudio graphics device  (use X11() device within RStudio). 


# Adding Windows Fonts 
R recognizes only a few basic fonts by default to use in plotting. Additional Windows true-type fonts (TTF) can be added as needed, once they have been installed.  The R windows device must also be instructed where to look for the fonts.

For a full description of how to add windows fonts see [Zev Ross blog](http://zevross.com/blog/2014/07/30/tired-of-using-helvetica-in-your-r-graphics-heres-how-to-use-the-fonts-you-like-2/).

Follow these basic steps:  

  1.  Create an R font database table (fonttable) with the following command: `extrafont::font_import()`  
      *  you only need to run this once, but must re-run it if you install more fonts to windows.  
      *  it takes awhile, so you don't want to re-run frequently.  
  2.  Register the fonts with the Windows device:  `extrafont::loadfonts(device = "win", quiet = T)`  
      *  Only need to do this once per R session  
  3.  To change the text, make sure to reference the font family name correctly.  you may need to reference the fonttable:  
      *  create a table and view it using the `View(extrafont::fonttable())` command  
      *  the `FontName` variable contains the name you need  
  4.  PDF/Postscript only. Ghostscript must be installed to embed fonts into a PDF file or Postscript file  
      *  Install Ghostscript by downloading from [their site](https://ghostscript.com/download/gsdnld.html)  
      *  Tell R where Ghostscript exe file is located: `Sys.setenv(R_GSCMD="C:/Program Files/gs/gs9.20")`  
      *  create a ggplot or other plot as a PDF:  `ggsave("newfont.pdf")`  
      *  Embed the fonts:  `embed_fonts("newfont.pdf")`  

At first, I though this was not working- but I was not being careful enough with the exact "FontName" variable. If you have trouble, doublecheck.      
```{r}
#windowsFonts("Cabin-Regular" = windowsFont("Cabin"))
extrafont::loadfonts(device = "win", quiet = T)
# View(extrafont::fonttable())
# Data for plotting
  a <- rnorm(500, 6, 1)
  b <- rnorm(400, 5, 1.5)
  c <- c(rnorm(200, 3, .7), rnorm(50, 7, 0.4))
  d <- c(rnorm(200, 2, 0.7), rnorm(300, 5.5, 0.4), rnorm(100, 8, 0.4))
df <- data.frame(
  "Distribution" = c(rep("Unimodal 1", length(a)),
                     rep("Unimodal 2", length(b)),
                     rep("Bimodal", length(c)),
                     rep("Trimodal", length(d))),
  "Value" = c(a, b, c, d))
df$Distribution <- factor(df$Distribution,
                          levels(df$Distribution)[c(3, 4, 1, 2)])
ggplot(df, aes(Distribution, Value)) +
  geom_violin(fill="steelblue", alpha=0.2) +
  geom_sina(scale=FALSE) +
  annotate(geom="text", x=3, y=9, label= "\"Annie Use Your Telescope\"\nFont", family="Annie Use Your Telescope", 
           lineheight=.8, color="red") +
  annotate(geom="text", x=3, y=0, label= "\"Roboto Condensed\" Font", family="Roboto Condensed", 
           lineheight=.8, color="red") +
  annotate(geom="text", x=1, y=1.75, label="\"Bauhaus 93\"\nAnnotation text", colour="red",
             size=7, family="Bauhaus 93", angle=30)

```

# Adding Google Fonts

Example on how to pull in fonts from GoogleFonts and create a family name in R: 
```{r, message=FALSE, warning=FALSE}
font.add.google("Gochi Hand", "gochi")
font.add.google("Schoolbell", "bell")
font.add.google("Fredericka the great", "frederickaTG")
font.add.google("Annie Use Your Telescope", "annie")

```


## `showtext` and GoogleFonts  

Showtext converts the text to a raster image and then adds it to the plot.

### Nonstandard Fonts in base plot

The following example adds different fonts to various text elements in ggplot:
(note: works only when rendering to X11 output within RStudio).  

Note: to make this work with RMD -> html output, add `fig.showtext=TRUE` to the chunk header. This automatically invokes the `showtext.auto()` command, which would otherwise need to be run in an interactive session.

```{r, echo=TRUE, fig.showtext=TRUE, message=FALSE, warning=FALSE}
library(showtext)
set.seed(123)
data<-data.frame(x=rnorm(1000))
## Manually open a graphics device if you run this code in RStudio
#X11({ # use this if plotting in RStudio; open font-friendly device to plot in 
#showtext.auto() ## Automatically use showtext to render text
hist(data$x, col="steelblue", breaks=50, border="white", xlab="", ylab="", main="")
title("Histogram of Normal Random Numbers", cex.main=4, family="annie")
  title(xlab = "Distribution of x", family = "gochi", cex.lab = 3)
  title(ylab="Density", family = "bell", cex.lab = 3)
#})
```


### Nonstandard Fonts in `ggplot`

The following example adds different fonts to various text elements in ggplot:
```{r, echo=TRUE, fig.showtext=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(formattable)
library(htmltools)
## Manually open a graphics device if you run this code in RStudio
#x11() # use this if plotting in RStudio; open font-friendly device to plot in 
showtext.auto() ## Automatically use showtext to render text
ggplot(data, aes(x)) + geom_density(fill = "steelblue", alpha=0.75) + 
  labs(title="Histogram of Normal Random Numbers",
       x="Distribution of x",
       y="Density") +
  theme_bw() + 
  theme(title = element_text(family="frederickaTG", color="steelblue", size = 32),
        axis.text.x = element_text(family="frederickaTG", color="steelblue", size = 18),
        axis.text.y = element_text(family="frederickaTG", color="steelblue", size = 18))        
#  theme(title=element_text(family="annie")) 
```

## `extrafont` and TTF  

This is useful for registering and using system fonts in PDF or postscript files.


```{r}
library(extrafont)
# font_import()
```



# Sparklines  
```{r}
library(sparkline)
x = rnorm(20)
y = rnorm(20)
```


[Sparklines](https://github.com/htmlwidgets/sparkline) is a package which allows addition of inline or table sparklines in html output from Rmd documents.  

## Examples:
Note: sparklines are interactive in html documents. Mouse over the data points.
an inline Sparkline: `r sparkline(x)`  
an inline barplot with absolute `r sparkline(abs(x), type = 'bar')` or including negative values  `r sparkline(x, type = 'bar')`  
an inline boxplot `r sparkline(x, type = 'box')`  

An interesting use of Sparklines is to add a visual representation of the data to a summary table:

Example 1 is using manually formatted table (inline text):  

| Stock | Sparkline         | Boxplot  
|-------|-------------------|--------
| x     | `r sparkline(x)`  | `r sparkline(x, type ='box')`
| y     | `r sparkline(y)`  | `r sparkline(y, type ='box')`

[Example 2 from StackOverlow](https://github.com/renkun-ken/formattable/issues/10#issuecomment-110864675) is using the `formattable` package for table printing,  from a code chunk:    
```{r}
df = data.frame(id=c("x","y"),"sparkline"=NA, "sparkbox" = NA)

formattable(
  df,
  list(
    "sparkline"=function(z){
      sapply(
        paste0(text='`r sparkline(',df$id,')`')
        ,function(md) knitr::knit(text=md, quiet=T)
      )
    }
    ,sparkbox =function(z){
      sapply(
        paste0(text='`r sparkline(',df$id,',type="box")`')
        ,function(md) knitr::knit(text=md, quiet=T)
      )
    }
  )
)
```

[Another example from StackOverflow](https://github.com/renkun-ken/formattable/issues/10#issuecomment-110864675)  
Example works in RStudio session but not in html output


An html table with color-bar in a code block from [stack overflow](http://stackoverflow.com/questions/32841221/add-sparkline-graph-to-a-table):
```{r cars}
library(dplyr)
library(formattable)
res <- 
  iris %>% 
  group_by(Species) %>% 
  summarise(N=n(),
            SL_Mean=round(mean(Sepal.Length),3),
            SL_SD=round(sd(Sepal.Length),3),
            SW_Mean=round(mean(Sepal.Width),3),
            SW_SD=round(sd(Sepal.Width),3))

#using formattable
formattable(res,
            list(
              SL_Mean=color_bar("pink", "proportion"),
              SL_SD=color_bar("steelblue")))
```

Some details from [leonawicz GitHub blog](http://leonawicz.github.io/HtmlWidgetExamples/ex_dt_sparkline.html) provides some details on making very nice tables from Rmarkdown to html, although it includes a lot of extra javascript that may not be needed for basic tables.

The [`sparktable` package](https://github.com/alexkowa/sparkTable) provides tools for making Tables which contain sparkline objects. Inline sparklines can be created using `sparktable::newSparkLine()`, but the result must be saved as an image, and then referenced inline. The result is also not interactive.  

The best reference for this package is the [R Journal manuscript](https://journal.r-project.org/archive/2015-1/templ-kowarik-meindl.pdf)  
```{r}

```


