---
title: "Graphics and Text"
author: "JML"
date: "January 12, 2017"
output: html_document
---

# Fonts

## `sysfont` package  


## `extrafont` package  
Very useful for TTF in PDF

## `showtext` package  
Flexible for non-standard fonts in R graphs
Doesn't work with RStudio graphics device  (use X11() device within RStudio)
Can pull in [Google fonts](https://fonts.google.com/)  

Example on how to pull in fonts from GoogleFonts and create a family name in R: 
```{r}
library(sysfonts)
font.add.google("Gochi Hand", "gochi")
font.add.google("Schoolbell", "bell")
font.add.google("Fredericka the great", "frederickaTG")
font.add.google("Annie Use Your Telescope", "annie")

```

## Nonstandard Fonts in base plot

The following example adds different fonts to various text elements in ggplot:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(showtext)
set.seed(123)
data<-data.frame(x=rnorm(1000))
## Manually open a graphics device if you run this code in RStudio
x11() # use this if plotting in RStudio; open font-friendly device to plot in 
showtext.auto() ## Automatically use showtext to render text
hist(data$x, col="steelblue", breaks=50, border="white", xlab="", ylab="", main="")
  title("Histogram of Normal Random Numbers", family="frederickaTG", plot.new=F)
  title(xlab = "Distribution of x", family = "gochi", cex.lab = 2)
  title(ylab="Density", family = "bell", cex = 2.5)
```


## Nonstandard Fonts in `ggplot`

The following example adds different fonts to various text elements in ggplot:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(showtext)
library(ggplot2)

set.seed(123)
data<-data.frame(x=rnorm(1000))
## Manually open a graphics device if you run this code in RStudio
#x11() # use this if plotting in RStudio; open font-friendly device to plot in 
showtext.auto() ## Automatically use showtext to render text
ggplot(data, aes(x)) + geom_density(fill = "steelblue", alpha=0.5) + 
  labs(title="Histogram of Normal Random Numbers",
       x="Distribution of x",
       y="Density") +
#  theme_bw() + 
  theme(plot.title=element_text(family="annie"),
        axis.title.y=element_text(family="bell"),
        axis.title.x=element_text(family="frederickaTG")) 
#title(ylab = "Frequency", family = "gochi", cex.lab = 2)
#text(2, 70, "N = 1000", family = "bell", cex = 2.5)
```

