---
title: "purrr Notes"
Date: "`r Sys.Date()`"
output: github_document
---
load common packages first:  
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
```

Common tasks can be simplified using `apply` functions rather than "for" loops, as would be used in other programming languages.  The `purrr` package includes the "tidyverse" versions as the `map` functions.

# Basics  

# Examples  

# My Tasks  

## Data Collection  

My research team commonly uses data collection into a standardized excel file (yes, really), with a standardized data summary for each study day.  Data from multiple sheets can be extracted simultaneously for a whole study using the `purr` functions.

Collect the file names from the data directory, including all subdirectories:  
note, this assumes that your data directory is located in the working directory.
```{r}
# get file names to import
#fileNames <- 
  dir(file.path(getwd(),"datasheets"), pattern = ".xlsx", recursive = TRUE)
  dir(pattern = ".xlsx", recursive = TRUE)
```

Create a dataFrame with ID, Date, and a list-column containing the data from each of the datasheets.  
note, pulling the id and date from standard format of the file names.  
```{r}
# Create a DataFrame
# and import the data needed for "minmod" 
data <- data_frame(
  id = gsub(".*/([0-9]*)_.*", "\\1", fileNames), # the id for each
  date = gsub(".*/[0-9]*_([0-9]*-[0-9]*-[0-9]*).*", "\\1", fileNames), # the date for each
  filename = fileNames) %>% mutate(
    ivgtt = map(paste0(getwd(),"/datasheets/", fileNames), read_excel, sheet="to minmod"))
```

View the data. The `ivgtt` data from each sheet is stored as a list-column.:  
```{r}
data
```

The data can be extracted into "long" format using the `unnest` command.  
A sample of the result is below.
```{r}
data %>% unnest() %>% 
  select(-filename) %>% 
  filter(time<5)
```

Plot some results:
```{r}
data %>% unnest() %>% group_by(id, date) %>%  
  ggplot(aes(x=time, group=interaction(id, date))) +
  geom_line(aes(y=insulin), color="steelblue") +
  geom_line(aes(y=glucose), color="red") +
  facet_wrap(~id) + theme_bw()
  

```

